---
title: "Play with PLOTLY R package"
author: "Qian Feng"
date: 2018-03-29
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->
I came across one website introducing [plotly](https://plotly-book.cpsievert.me) R package, and was addicted  for a few days, just share some of these amazing plots with you. 
The main scheme is introduce three basic examples, followed by scatter plot, line plot and ribbons.I will conclude this section with two great animation examples.

```{r, message=FALSE, warning=FALSE}
##install.packages("plotly")
##packageVersion("plotly")
library(plotly)
library(ggplot2)
```

## Basic example 1
Interface modified by layout() function:"click + drag"event, rangeslider() will add a slider under this figure.

```{r, message=FALSE, warning=FALSE}
library(forecast)
p <- ggplot(fortify(gold), aes(x, y)) + geom_line()
gg <- ggplotly(p)
layout(gg, dragmode = "pan")
##rangeslider(gg)
```


## Basic example 2
Look at and learn the basic structre , firstly use ggplotly(), then %>% link several matrics.
```{r, message=FALSE, warning=FALSE}
p <- ggplot(MASS::geyser, aes(x = waiting, y = duration)) +
  geom_density2d()
ggplotly(p, originalData = FALSE) %>% 
  group_by(piece) %>%
  slice(which.min(y)) %>% 
  add_text(
    text = ~level, size = I(10), color = I("black"), hoverinfo = "none"
  )
```

## Basic example 3
```{r, message=FALSE, warning=FALSE}
p <- ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() + geom_smooth()
style(p, hoverinfo = "none", traces = 2:3)
```

Of course, we even could add annotations in detailed positions by function add_annotations() and add_segments().
```{r, message=FALSE, warning=FALSE}
p <- ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() + geom_smooth()
p %>%
  ggplotly(layerData = 2, originalData = FALSE) %>%
  add_fun(function(p) {
    p %>% slice(which.max(se)) %>%
      add_segments(x = ~x, xend = ~x, y = ~ymin, yend = ~ymax) %>%
      add_annotations("Maximum uncertainty", ax = 60)
  }) %>%
  add_fun(function(p) {
    p %>% slice(which.min(se)) %>%
      add_segments(x = ~x, xend = ~x, y = ~ymin, yend = ~ymax) %>%
      add_annotations("Minimum uncertainty")
  })
```
## 2D and 3D Scatter plot 
This time we use plot_ly(), previous sections are all employed ggplotly().

```{r,message=FALSE, warning=FALSE}
##I(2)is triangle, I(3)is +
subplot(
  plot_ly(x = 1:25, y = 1:25, symbol = I(1:25), name = "pch"),
  plot_ly(mpg, x = ~cty, y = ~hwy) %>% 
    add_markers(symbol = I(1), color = I("red"),name = "red")
)

```

```{r}
p <- plot_ly(mpg, x = ~cty, y = ~hwy, alpha = 0.5)
subplot(
  add_markers(p, color = ~cyl, showlegend = F,name = "mpg") %>% 
    colorbar(title = "Viridis")
)
```

Last ones are two 3D examples, you are free to rotate the figures.
```{r}
plot_ly(mpg, x = ~cty, y = ~hwy, z = ~cyl) %>%
  add_markers(color = ~cyl)
```

```{r}
x <- seq_len(nrow(volcano)) + 100
y <- seq_len(ncol(volcano)) + 500
plot_ly() %>% add_surface(x = ~x, y = ~y, z = ~volcano)
```

A powerful functiom to visualize data iris:
```{r,message=FALSE, warning=FALSE}
pm <- GGally::ggpairs(iris)
ggplotly(pm)
```

## Line plots
Lower image are the same data with upper image, just to compare the effects of add_lines(p, color = ~city) and add_lines(p, linetype = ~city).
```{r,message=FALSE, warning=FALSE}
library(dplyr)
top5 <- txhousing %>%
  group_by(city) %>%
  summarise(m = mean(sales, na.rm = TRUE)) %>%
  arrange(desc(m)) %>%
  top_n(5)

p <- semi_join(txhousing, top5, by = "city") %>%
  plot_ly(x = ~date, y = ~median)
subplot(
  add_lines(p, color = ~city),
  add_lines(p, linetype = ~city),
  shareX = TRUE, nrows = 2
)
```

## Ribbons
```{r,message=FALSE, warning=FALSE}
library(broom)
m <- lm(mpg ~ wt, data = mtcars)
broom::augment(m) %>%
  plot_ly(x = ~wt, showlegend = FALSE) %>%
  add_markers(y = ~mpg, color = I("red")) %>%
  add_ribbons(ymin = ~.fitted - 1.96 * .se.fit, 
              ymax = ~.fitted + 1.96 * .se.fit, color = I("gray80")) %>%
  add_lines(y = ~.fitted, color = I("steelblue"))
```

## Animation 1
Return to use function ggplotly()
```{r,message=FALSE, warning=FALSE}
data(gapminder, package = "gapminder")
head(gapminder)
gg <- ggplot(gapminder, aes(gdpPercap, lifeExp, color = continent)) +
  geom_point(aes(size = pop, frame = year, ids = country)) +
  scale_x_log10()
ggplotly(gg)
```

## Animation 2

```{r}
meanLife <- with(gapminder, tapply(lifeExp, INDEX = continent, mean))
gapminder$continent <- factor(
  gapminder$continent, levels = names(sort(meanLife))
)
base <- gapminder %>%
  plot_ly(x = ~gdpPercap, y = ~lifeExp, size = ~pop, 
          text = ~country, hoverinfo = "text") %>%
  layout(xaxis = list(type = "log"))
base %>%
  add_markers(data = gapminder, frame = ~continent) %>%
  hide_legend() %>%
  animation_opts(frame = 1000, transition = 0, redraw = FALSE)
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
